---
title: "Northdale model contrast threat and empath"
author: "CGT"
date: "Wednesday, July 30, 2014"
output: html_document
---

This document records the re-analysis of a data file sent to me by Kevin D and 
Mike Q.  I am using LAVAAN in R to model Prejudice as some function of Threat, Empathy, and Closeness, with SEM methods.

#   Read the data
#   read tab-delimited data file:
```{r}
# setwd("d:/Dropbox/research/Social Psychology/northdale/Tradeoff threat vs empathy/")
data<-read.table("Northdale survey (reverse coded).dat",
                    header = TRUE,sep="\t")
library(lavaan);library(knitr);library(psych);library(GPArotation);library(dplyr)
```


#   scale variables.

We want to scale all the variables so they are on 1-5 scale.  We also compute alpha and omega.

```{r}
#   scale distance
    data$crowflies_address_infstlmnt<-(1-data$crowflies_address_infstlmnt/max(data$crowflies_address_infstlmnt,na.rm=T))
    data$Cnt_IS_see<-data$Cnt_IS_see
    data$Cnt_IS_smell<-data$Cnt_IS_smell
    data$Cnt_IS_hear<-data$Cnt_IS_hear
    data$closeness_scale<-(data$Cnt_IS_see + data$Cnt_IS_smell + data$Cnt_IS_hear + data$crowflies_address_infstlmnt)
    closenessdf<-data.frame(data$Cnt_IS_see,data$Cnt_IS_smell,data$Cnt_IS_hear,data$crowflies_address_infstlmnt)    
    omega(closenessdf,nfactors=1)
  

#   Threat
    data$Thrt_Phy1<-data$Thrt_Phy1/3
    data$Thrt_Phy2<-data$Thrt_Phy2/3
    data$Thrt_Phy3<-data$Thrt_Phy3/3
    data$threat_scale<-data$Thrt_Phy1+data$Thrt_Phy2+data$Thrt_Phy3
    threatdf<-data.frame(data$Thrt_Phy1, data$Thrt_Phy2, data$Thrt_Phy3)    
    omega(threatdf,nfactors=1)
#   so what 


  
#   Empathy
    data$Empath1<-data$Empath1/3    
    data$Empath2<-data$Empath2/3
    data$Empath3<-data$Empath3/3
    data$empathy_scale<-(data$Empath1 + data$Empath2 + data$Empath3)
    empathydf<-data.frame(data$Empath1 ,  data$Empath2 ,  data$Empath3)    
    alpha(empathydf)
    omega(empathydf,nfactors=1)



#   Prejudice
    data$Att_IS1<-(data$Att_IS1-1)/6.75    
    data$Att_IS2<-(data$Att_IS2-1)/6.75  
    data$Att_IS4<-(data$Att_IS4-1)/6.75  
    data$prejudice_scale<-(data$Att_IS1 + data$Att_IS2 + data$Att_IS4)
    prejudicedf<-data.frame(data$Att_IS1 ,  data$Att_IS2 ,  data$Att_IS4)    
    omega(prejudicedf,nfactors=1)
    alpha(prejudicedf)


#   Contact quantity
    data$Cnt_pos_B<-data$Cnt_pos_B/2
    data$Cnt_pos_IS1<-data$Cnt_pos_IS1/2
    data$contactquant_scale<-data$Cnt_pos_B+data$Cnt_pos_IS1
    contactquantdf<-data.frame(data$Cnt_pos_B,data$Cnt_pos_IS1)    
    omega(contactquantdf,nfactors=1)
    alpha(contactquantdf)



#   Contact_avoid
    data$cnt_avd2<-data$cnt_avd2/3
    data$cnt_avd3<-data$cnt_avd3/3
    data$cnt_avd4<-data$cnt_avd4/3
    data$contactavoid_scale<-data$cnt_avd2+data$cnt_avd3+data$cnt_avd4
    contactavoiddf<-data.frame(data$cnt_avd2,data$cnt_avd3,data$cnt_avd4)    
    omega(contactavoiddf,nfactors=1)
    alpha(contactavoiddf)


#   Contact_quality
    data$Cnt_Qul_IS1<-data$Cnt_Qul_IS1/3
    data$Cnt_Qul_IS2<-data$Cnt_Qul_IS2/3
    data$Cnt_Qul_IS3<-data$Cnt_Qul_IS3/3
    data$contactqual_scale<-data$Cnt_Qul_IS1+data$Cnt_Qul_IS2+data$Cnt_Qul_IS3
    contactqualdf<-data.frame(data$Cnt_Qul_IS1,data$Cnt_Qul_IS2,data$Cnt_Qul_IS3)    
    omega(contactqualdf,nfactors=1)
    alpha(contactqualdf)

#   Let's get some descriptive statistics on the scale variables
    library(dplyr)
    datax<-tbl_df(data) # convert to format dplyr likes
    statsum<-summarise_each(datax,funs(mean(.,na.rm=T),sd(.,na.rm=T),max(.,na.rm=T)),matches("scale"))
    print(statsum)
    
#   Let's take a look at the inter-connectedness of the contact scales
    contactvars<-data.frame(data$contactqual_scale,data$contactquant_scale,data$contactavoid_scale,data$closeness_scale)
    pairs.panels(contactvars)
    princomp(~.,contactvars)
    principal(contactvars,2,rotate="cluster")

    # The analysis above shows that the contact vars are correlated, and that there appear to be
    # two separate factors, strongly defined by contactquant and contactavoid, with contactqual
    # loading on both.  It would be useful to test a CFA with original items, proposing 3 separate

    contactfactors<-"

 
    #   Contact_quant
        contactquant=~Cnt_pos_B+Cnt_pos_IS1

    #   Contact_avoid
        contactavoid=~cnt_avd2+cnt_avd3+cnt_avd4

    #   Contact_quality
        contactqual=~Cnt_Qul_IS1+Cnt_Qul_IS2+Cnt_Qul_IS3

"

      #     We now run the CFA
            contactfactorsfit<-sem(contactfactors,data=data)
            kable(summary(contactfactorsfit,rsquare=T),format="html")            
            parameterEstimates(contactfactorsfit) 
            resid(contactfactorsfit, type = "standardized")
            fitmeasures(contactfactorsfit)
            modindices(contactfactorsfit)


#   Kevin wants to take a look at the negative contact variables
    contactnegativedf<-data.frame(data$Cnt_neg_IS1,data$Cnt_neg_IS2,data$Cnt_neg_IS3)
    pairs.panels(contactnegativedf)
    data$contactnegative_scale<-data$Cnt_neg_IS1 + data$Cnt_neg_IS2 + data$Cnt_neg_IS3    

    # clearly good properties, so we add this to the contactvars dataframe, and look again
    contactvars<-data.frame(data$contactqual_scale,data$contactquant_scale,data$contactavoid_scale,data$closeness_scale,data$contactnegative_scale)
    pairs.panels(contactvars)
    princomp(~.,contactvars)
    principal(contactvars,2,rotate="cluster")
    
    # Very clear pattern, again.  Now take a look at negative emotions
    datax<-tbl_df(data)
    negativeemotionsdf<-select(datax,starts_with("Emot_IS_G"))
    negativeemotionsdf<-negativeemotionsdf[,1:8]
    pairs.panels(negativeemotionsdf)
    princomp(~.,negativeemotionsdf)
    principal(negativeemotionsdf,2,rotate="cluster")


```


We now define a LAVAAN model (measurement + structural).  This model proposes 
that there are two paths to prejudice, one through empathy (a classic 
theoretical proposition in this area), and another through threat, 
which itself is a function of closeness or distance of the respondents 
to settlers. But we now complexify the model a bit, getting contact to
predict empathy.


```{r}

medcloseness2<-"
        
#   MEASUREMENT MODEL (Lavaan fixes first items for scale to 1.0, 
repeated here to be clear)

#   Closeness
    closeness=~1.0*(crowflies_address_infstlmnt) + Cnt_IS_see + Cnt_IS_smell+ Cnt_IS_hear

#   Threat
    threatphy=~1.0*Thrt_Phy1 + Thrt_Phy2 + Thrt_Phy3

#   Empathy
    empathy=~1.0*Empath1 + Empath2 + Empath3

#   Prejudice
    prejudice=~1*Att_IS1 + Att_IS2 + Att_IS4
    
#   Contact_pos
#    contactpos=~Cnt_Qul_IS1+Cnt_Qul_IS2+Cnt_Qul_IS3

    contactquant=~Cnt_pos_B+Cnt_pos_IS1

#   Contact_avoid
    contactavoid=~cnt_avd2+cnt_avd3+cnt_avd4

#   Contact_quality
    contactqual=~Cnt_Qul_IS1+Cnt_Qul_IS2+Cnt_Qul_IS3

#   Negative_emotions
#    negemotion=~Emot_Is_G2+Emot_Is_G4+Emot_Is_G5

#   Negative contact
    contactnegative=~Cnt_neg_IS1 + Cnt_neg_IS2


#   REGRESSIONS / STRUCTURAL MODEL
#   contactqual~closeness+contactnegative
    contactquant~contactqual+contactnegative
    contactqual~threatphy+contactnegative
    contactnegative~closeness
    threatphy~closeness+contactnegative
    empathy~contactqual+threatphy
    contactavoid~threatphy+contactqual+contactnegative
#   contactavoid~closeness
#   contactavoid~threatphy
#   negemotion~contactquant+threatphy
    prejudice~contactquant+contactavoid
 
       
"

  # test



```

We now run the SEM, asking for a summary of the model, parameter estimates, standardized residuals, comprehensive reporting on measures of fit, and modification indices.  Standard SEM analysis.
Cannot get Git to update though

```{r}

medcloseness2fit<-sem(medcloseness2,data=data)
kable(summary(medcloseness2fit,rsquare=T),format="html")            
parameterEstimates(medcloseness2fit) 
resid(medcloseness2fit, type = "standardized")
fitmeasures(medcloseness2fit)
modindices(medcloseness2fit)

```


We do some additional exploration.  THe questions now are whether Closeness moderates the relationship between 
contact and prejudice

```{r echo=FALSE}  

# check out scatterplots etc.
modeldatadf<-data.frame(data$closeness,data$contactquant,data$contactavoid,data$threat,data$empathy,data$prejudice)
pairs.panels(modeldatadf)

#  check out possible moderation of contact-prej relationship by closeness
closeness_X_contactprej<-lm(prejudice_scale~closeness_scale*contactquant_scale,data=data)
summary(closeness_X_contactprej)

#  check out possible moderation of contactqual-prej relationship by closeness
closeness_X_contactqualprej<-lm(prejudice_scale~closeness_scale*contactqual_scale,data=data)
summary(closeness_X_contactqualprej)


# Check out possible contact-quality dependence on closeness
scatter.smooth(data$closeness_scale,data$contactqual_scale)
scatter.smooth(data$contactqual_scale,data$contactquant_scale)
cor(data$contactqual_scale,data$contactquant_scale,use="pairwise.complete")

# Check out possible moderation of contactquant-prej by contact-qual
contactqual_X_contactquantprej<-lm(prejudice_scale~contactqual_scale*contactquant_scale,data=data)
summary(contactqual_X_contactquantprej)


```








